---
title: "GitHub Analysis"
author: "Zhian N. Kamvar"
date: "3/30/2020"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

The source for all the repositories is located at <https://carpentries.github.io/curriculum-feed/carpentries_lessons.json>. To 
avoid having to pull from it multiple times I'm going to set up a folder:

```{r packages}
library("fs")       # Filesystem navigation
library("jsonlite") # parsing JSON files
library("purrr")    # handling lists (JSON files)
library("dplyr")    # handling data frames and magic
library("ggplot2")  # visualization
library("forcats")  # ordering factors
library("magrittr") # for the %T>% pipe I love so well
library("here")     # so I can always remember where I started
```


```{r data}
raw_data<- here::here("data", "raw")
raw_lessons <- fs::path(raw_data, "carpentries_lessons.json")
if (fs::file_exists(raw_lessons)) {
  
} else {
  fs::dir_create(raw_data)
  res <- download.file(
    url = "https://carpentries.github.io/curriculum-feed/carpentries_lessons.json",
    destfile = raw_lessons
  )
  if (res == 0L) {
    fs::file_chmod(raw_lessons, "a-wx")
  }
}

lessons <- read_json(raw_lessons, simplifyVector = FALSE) %>%
  purrr::map_dfr(.f = dplyr::as_tibble) %>%
  dplyr::mutate_if(is.character, 
    ~dplyr::if_else(. == "NA", as.character(NA), .)) %>%
  dplyr::mutate(life_cycle = forcats::fct_relevel(life_cycle,
      "pre-alpha", "alpha", "beta", "stable", "on-hold", "deprecated"
    )) %>%
  dplyr::mutate(life_cycle = forcats::fct_explicit_na(life_cycle)) %>%
  print()
```

# First look at available lessons

I'm going to take a look at all of the software carpentry lessons first. There
are currently `r nrow(lessons)` available lessons, but they are in different
stages of completion:

```{r lesson vis}
ggplot(lessons, aes(y = life_cycle, fill = curriculum)) +
  geom_bar(orientation = "y")
```

Let's inspect the ones that are stable and then look at the ones that are lower
down, first we should filter for those that are stable and actually have GitHub
URLs:

```{r stable}
stable <- lessons %>% filter(
  life_cycle == "stable", 
  curriculum != "instructor-training",
  !is.na(curriculum), # these are template repositories
  !is.na(URL)         # no GitHub URL is not very useful for me
)
stable
```

I end up with `r nrow(stable)` repositories to play with:

```{r curriculum-language}
ggplot(stable, aes(y = curriculum, fill = language)) +
  geom_bar(orientation = "y")
```

# Inspection of repository features

I will use the {gh} package to inspect the features of each repository:

 - tags
 - directory structure
 - dependencies

# Session Information

```{r sessioninfo}
sessioninfo::session_info()
```

